---
title: Handling Large Data in R, the Naive Way
author: Siqi Zhang
date: '2019-07-16'
slug: naive-caching-r-objects-onto-disk
categories:
  - R
tags: []
---

Recently I ran accross a problem: The program I'm writing reads from hundreds of large data sets totaling in tens of gigabytes, and the memory on my or any potential user's machine simply won't fit. 

Ideally, for caching objects on the disk, I would have wanted to use more sophisticated solusions such as Roger Peng's [filehash](https://cran.r-project.org/web/packages/filehash/index.html). Instead, I just wanted to simply dump everything onto the disk, and read & process each as needed, with the following solution: 

- Write each data set as a binary file to the disk
- Make a function that reads from its own binary file when called

It goes as follows:
```{r}
stash <- function(object, dir_path = tempdir()){
        file_name <- paste0(paste0(sample(c(letters, LETTERS, 0:9), 20, TRUE), collapse = ""), ".RStash")
        file_path <- file.path(dir_path, file_name)
        saveRDS(object, file_path)
        f <- function(){
                if (!file.exists(file_path)){
                        stop("stash file missing.")
                } else {
                        readRDS(file_path)
                }
        }

        structure(f,
                  class = c("stash_pointer", class(f)),
                  file_path = file_path,
                  obj_size = format(object.size(object), unit = "MB", digits = 2),
                  obj_class = class(object)
                  )
}


```


So calling `stash()` saves a binary of the object onto the disk, and returns a function that is going to read from this the file when called, with relevant metadata. I just call the cache on the disk "stash" and this function a "stash pointer".  


Let's test it out.

```{r}
mtcars2 <- stash(mtcars)
mtcars2() 
```
Perfect! 


Let's add some housekeeping funcitions, too.
```{r}
print.stash_pointer <- function(x){
        cat(paste0("<stash_pointer>", " `", attr(x, "obj_class")[1],"` ", attr(x, "obj_size"), "\n"))
        cat("- ", paste(attr(x, "file_path")))
}

# delete the cache on disk
clear_stash <- function(stash_pointer){
        file_path <- attr(stash_pointer, "file_path")
        if (file.exists(file_path)){
                file.remove(file_path)      
        } 
}

stash_exists <- function(stash_pointer){
        file.exists(attr(stash_pointer, "file_path"))
}

is_stash_pointer <- function(x){
        inherits(x, "stash_pointer")
}
```



```{r}
mtcars2
```

```{r}
is_stash_pointer(mtcars2)
```

```{r}
clear_stash(mtcars2)
```

```{r}
stash_exists(mtcars2)
```

```{r}
df_list <- list(mtcars, iris, chickwts, PlantGrowth, USArrests)
```

Let's utilize `purrr`'s functional programming interface and `stash()` everything in `df_list`:
```{r}
require(purrr)
stash_pointer_list <- map(df_list, stash)
```

The result is a list of `stash_pointer`s:
```{r}
stash_pointer_list
```

See the column name of every data frame:
```{r}
map(stash_pointer_list, ~ exec(.) %>% colnames())
```

Cheers! 
