---
title: "Variadic function in R and Python"
author: "Siqi Zhang"
date: '2019-03-29'
slug: variadic-function-in-r-and-python
tags:
- Programming
- function
- variadic
categories: [R, Python]
---
```{r, setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```


In R, it is most usual to see all arguments defined explicitly for a funciton, such as in:

```{r, eval = FALSE}
rnorm(n, mean = 0, sd = 1)
```
```{r}
rnorm(2)
```

Other times, the function may take any number of arguments, such as in: 

```{r, eval = FALSE}
sum(..., na.rm = FALSE)
```
```{r}
sum(1,2,3, c(4,5,6), 7:9)
```

Functions like `sum()` are called [_variadic functions_](https://en.wikipedia.org/wiki/Variadic_function), where `...` serves as the placeholder for the unspecified arguments. Note that the correct name of `...`, according to the offical [R Language Definition](https://cran.r-project.org/doc/manuals/R-lang.html#Dot_002ddot_002ddot), is _dot-dot-dot_, not _eclipse_. It is not unusual to see smart people [make this little  mistake](https://ipub.com/r-three-dots-ellipsis/). 

The `...` can also take arguments The following code calls `c()` function, and uses argument names as the element names in the resulting vector.
```{r}
c(foo = TRUE, bar = FALSE)
```


### Writing a variadic function
You may decide on whether to use variadic arguments when writing your own function.

The following function takes an `x` and an `y` as arguments, and results the sum of `x` and `y` and 7. You must specify `x` and `y`, and only `x` and `y`, when calling this funciton.

```{r}
add_seven <- function(x, y){
        x + y + 7
}

add_seven(x = 10, y = 10)
add_seven(10, 10)
```


This following function uses `...` to take one, two or many numbers. There can only be one `...` in one function call; and the `...` stays the same through the whole call stack. That means you can pass along the `...` to another function that also takes `...`. 

```{r}
add_seven_variadic <- function(...){
        sum(..., 7)
}

add_seven_variadic()
add_seven_variadic(10)
add_seven_variadic(10, 10)
add_seven_variadic(10, 10, 10)
add_seven_variadic(x = 10, y = 10, z = 10) # argument names ignored here
```


You may also refer to the  `...` more than one time in the function body. The following function adds the `...` with 7, and then itself again. 

```{r}
add_seven_variadic_twice <- function(...){
        add_seven_variadic(...) + sum(...)
}

add_seven_variadic_twice(10)
add_seven_variadic_twice(10, 10)
```


### Capturing dot-dot-dot
It is also possible to _*capture*_ the `...` as a list. This is very helpful in advanced programming. The easiest way is just using `list(...)`.

In the following example, `tell_pocket()` captures the `...` as a list, and prints some messages accordingly.

```{r}

tell_pocket <- function(...) {
        dots <- list(...)
        
        cat("I have: \n")
        for (i in seq_along(dots)){
                cat("  ", names(dots)[i], ": ", dots[[i]], 
                    "\n", sep = "")
        }
}

coins <- 2.25
bills = 5

tell_pocket(
        cellphone = "Nokia 3310",
        money = coins + bills
)



```


### Dot Expansion

Using `list(...)` automatically evaluates any expresion in the `...`, this is called _dot expansion_.  

In our previous example, the `money` argument took the expression `coins + bills`, which is evaluated in the global environment, where `tell_pocket()` is called; the `cellphone` argument receives the string `"Nokia 3310"`, which will always evaluate to itself. 

However, we sometimes may wish to evaluate the arguments later, in a different context, or deparsing them into characters. In this case, we must avoid dot expansion and keeping them as-is. One way to do it is using `match.call(expand.dots = FALSE)$...`

In this example with `tell_pocket2`, we get values separately form the `left_pocket` and the `right_pocket`:

```{r}
tell_pocket2 <- function(pocket, ...) {
        dots <- match.call(expand.dots = FALSE)$...
        
        cat("I have: \n")
        for (i in seq_along(dots)){
                dot <- eval(dots[[i]], envir = pocket)
                cat("  ", names(dots)[i], ": ", dot, 
                    "\n", sep = "")
        }
        cat("in the", deparse(substitute(pocket)), "\n\n")
}

left_pocket <- list(
        coins = 0.75,
        bills = 25
)

right_pocket <- list(
        coins = 2.25,
        bills = 10
)

tell_pocket2(left_pocket, money = coins + bills)
tell_pocket2(left_pocket, coins = coins, bills = bills)
tell_pocket2(right_pocket, money = coins + bills)
```

The `tell_pocket2()` function used `eval()` to evaluate the expression in a specified environment, and used `substitute()` to capture the name supplied to an argument; these features fall into a grander scheme called _Non-standard Evaluation_(NSE) in R, which can be traced to R's [Lisp roots](https://www.stat.auckland.ac.nz/~ihaka/downloads/Compstat-2008.pdf
), and is far more sophisticated than that in Python. You may read more about NSE in [this chapter](http://adv-r.had.co.nz/Computing-on-the-language.html) from _Advanced R_ by Wickham.


### With Python
In Python, creating and using variadic functions is a little different. One uses `*args` and `**kwargs` to denote variadic arguments in place of `...`, meaning _arguments_ and _keyword arguments_. While it is legal to an arbitrary name succeding `*` or `**`, such as `*foo` and `**bar`, the previous example is the prevailing convention. 

There is no need to _capture_ the arguments, as `*args` and `**kwargs` are already `list` and `dictionary`, respectively.

Here are some examples in Python corresponding to previous R examples:
```{python}
def add_seven(x, y):
    return sum([x, y, 7])

print(add_seven(10, 10))
```

```{python}
def add_seven_variadic(*args):
    return sum([*args, 7]) 
    
print(add_seven_variadic())
print(add_seven_variadic(10))
print(add_seven_variadic(10, 10))
print(add_seven_variadic(10, 10, 10))

four_tens = [10, 10, 10, 10]
print(add_seven_variadic(*four_tens)) 
```

Note that the `sum()` function in Python is not a variadic function itself, and must take an _iterable_ as its argument; in this case, a `list`. 


```{python}
def tell_pocket(**kwargs):
    print("I have:")
    for key, value in kwargs.items():
        print("  {}: {}".format(key, value))

coins = 0.5
bills = 5

tell_pocket(
    cellphone = "Motorola 300",
    money = coins + bills
)
```

Additionally, it is possible to use `*args` and `**kwargs` together in one function. But `*args` must occur before `**kwargs`.
```{python}
def print_all(*args, **kwargs):
    for arg in args:
        print(arg)
    for kwarg in kwargs.items():
        print(kwarg)

print_all(1,2,3, x = 10, y = 20, z = 30)
```

To learn more about variadic funcitons in Python, [this article](https://www.digitalocean.com/community/tutorials/how-to-use-args-and-kwargs-in-python-3) sums things up very well. 

